Are Delayed Issues Harder to Resolve?

### 2015-12-31 phaseDelay4Emse.tex
Updates from ShareLaTeX

\section{Definitions \& Hypotheses}
This paper uses the following definitions:
\bi
\item
The {\em delayed issue effect}:   it is {\em exponentially}  more {\em difficult} to resolve  issues in a software project, the {\em longer} they remain.
\item
 {\em Longer} time is defined as per  Boehm'81~\cite{Boehm81}; i.e. the gap between the   phases where   issues are introduced and resolved.
\item
A measure collected in phase ${1,.,i,..j}$ is 
{\em exponential} when  that
   measure at phase $j$   
   is larger than the sum of that measure in 
earlier phases $1 \le i < j$. 
\item
Issues are more {\em difficult}  
when their resolution takes more time or costs more  (e.g. needs expensive
debugging tools or the skills of expensive developers).
\ei
Note that we use  the  term ``delayed issue effect'' as generalization of the
more specific rule  ``requirements errors are hardest to fix''.
This generalization is valid since the rationale for the rule about requirements
is usually done as per McConnell~\cite{mcconnell01}; i.e. small upstream mistakes very
early in the system can cascade into huge problems later in the lifecycle.
That said, we prefer our more general term ``delayed issue effect'' since not
only might requirements errors cascade, so too might analysis errors, design errors, etc.

@@

{\bf Claim3: Delayed Issues not Harder to Resolve.}
 In our sample of 
 171 software  projects developed, we will show
{\em no evidence}  that, during development, delayed issues were exponentially harder to resolve the longer they were left
 in the software.

@@

This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access
the data used to make prior conclusions.
As an example of this, \fig{steck} shows one 2004 survey that reports an 
exponential delayed issue  effect for 
requirements issues in eight case studies. Note that we cannot verify
{\em any} of those results since the links in the references of
that  survey are all broken (``page not found'' errors).

@@

 All in all, the empirical results of this section seem insufficient to justify
 the strength of the beliefs documented in the previous section. Adding to our doubts are  several studies that report less-than exponential
 increase in the difficulties associated with making the changes associated with delayed
 issues. 
Clutterbuck et al.~\cite{Clutterbuck09} studied a 5-month effort by a small-to-medium enterprise team developing a 71KLOEC web interface to a database application to implement 18 change requests-- see Figure~\ref{fig:clutterbuck} (note that these were for new and changed user requirements, not defects). Clutterbuck et al. found the cost of change to be relatively flat until the later phases, with much of the effort spent in analysis of the change requests~\cite{Clutterbuck09}. Note that in this study, the effort increased by only 60\% (see the start and end of the curve in Figure~\ref{fig:clutterbuck}).

Another example of less-than exponential explosion in the difficulty associated with delayed issues comes from  Royce~\cite{Royce98}. He
studied  a million-line, safety-critical missile defense system
(see  Figure~\ref{fig:royce}). Design changes (including architecture changes) required approximately twice the effort of implementation and test changes, and the cost-to-fix in implementation and test phases increased slowly. Boehm~\cite{Boehm10} attributes this success to the  development process, which focused on removing architecture risk early in the development lifecycle.

@@

 One feature of the above results will become very important in the subsequent discussion. We note that that in all
the results that report a delayed issue effect, those effects have began to manifest by the end of that project.

@@

\begin{figure}[!t] 
 \renewcommand{\baselinestretch}{0.7}
 \scriptsize
\begin{center}
\begin{tabular}{r@{~}|r@{~}r@{~}r|@{~}l@{~}l|@{~}r@{~}l}   
  \multicolumn{1}{c}{~} &\multicolumn{3}{c}{Percentiles}&\multicolumn{2}{l}{~~~~~~Phase}\\
N & 25th & 50th & 75th &   injected & removed   & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline
\\
 48&   14&   30&   51&Before&DesignInspect &1.00 & **********  \\
 53&   13&   28&   49&Before&CodeReview &0.93 & **********  \\
 94&   22&   45&   80&Before&CodeInspect &1.50 & ***************  \\
153&   38&   77&  130&Before&UnitTest &2.56 & **************************  \\
127&   30&   65&  128&Before&IntTest &2.16 & **********************  \\
 92&   24&   56&   89&Before&SysTest &1.86 & *******************  \\

\\
 56&   14&   35&   57&Planning&ReqtsReview &1.00 & **********  \\
 46&   11&   24&   39&Planning&DesignInspect &0.68 & *******  \\
 32&   18&   33&   84&Planning&UnitTest &0.94 & **********  \\

\\
184&   45&   92&  155&Reqts&ReqtsReview &1.00 & **********  \\
189&   46&   93&  155&Reqts&ReqtsInspect &1.00 & **********  \\
 61&   14&   33&   55&Reqts&DesignReview &0.35 & ****  \\
102&   24&   49&   83&Reqts&DesignInspect &0.52 & ******  \\
 59&   18&   38&   60&Reqts&CodeInspect &0.40 & ****  \\
 91&   22&   49&  104&Reqts&UnitTest &0.52 & ******  \\
100&   33&   67&  120&Reqts&IntTest &0.72 & ********  \\
 78&   28&  122&  237&Reqts&SysTest &1.31 & **************  \\
\\
210&   51&  103&  166&Design&DesignReview &1.01 & ***********  \\
207&   50&  101&  156&Design&DesignInspect &1.00 & **********  \\
134&   32&   66&  116&Design&CodeReview &0.65 & *******  \\
163&   39&   79&  135&Design&CodeInspect &0.78 & ********  \\
238&   58&  119&  183&Design&UnitTest &1.17 & ************  \\
172&   42&   88&  158&Design&IntTest &0.87 & *********  \\
163&   40&   80&  161&Design&SysTest &0.79 & ********  \\
 80&   21&   45&   73&Design&AcceptTest &0.44 & *****  \\

\\
235&   57&  115&  177&Code&CodeReview &1.00 &  **********  \\
228&   56&  113&  176&Code&CodeInspect &0.99 & ********** \\
243&   59&  125&  206&Code&IntTest &1.08 &     *********** \\
191&   46&   95&  173&Code&SysTest &0.82 &     ******** \\
269&   66&  133&  206&Code&UnitTest &1.15 &    ***********  \\
130&   31&   65&  122&Code&AcceptTest &0.56 &  ******  \\
 \end{tabular}
\end{center}
\caption{Distribution of fix times seen in SEI TSP data. Left hand side bars
show scale up on median values.
}
\label{fig:raw}
\end{figure}

@@

In that figure, the data is split out according to issues that were fixed in phase $Y$ after
being introduced in phase $X$. The data is sub-divided into tables according to $X$;
i.e. according to {\em before, planning, requirements, design} or {\em  code}. 
The stars on the right-hand-side of that figure show the scale ups associated with the phase removal stages,
expressed as a ratio of the first item in that block (which, by definition, is 1.00 for
that first item).
If  this data exhibited a delayed issue effect,  we would expect those bars in each sub-division to exhibit an
exponential shape.

\subsubsection{Comment 1}
Figure~\ref{fig:raw}
 data does not support that   
{\em  requirements errors are the most expensive to fix}.
Note that   median time spent fixing errors from {\em requirements} across the life cycle was 92, 93,  49, 33, 39, 49, 67, 122 minutes\footnote{If this time seems long, then recall
that it includes the time required  to (a)~collect data and realize there is an error;
(b)~prepare a fix;  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).}.
This is:
\bi
\item
Clearly longer than median times seen {\em Before } and in {\em Planning}
(30,28,45,77,65,56,24,35,33);
\item
Somewhat shorter, but still not exponentially smaller,  than the median times see after {\em requirements} in {\em design}
(103,101, 79,66, 119,88, 80,45) and in {\em code} (65,115, 113,125, 95,133). 
\ei
For completeness, we ran an A12 test comparing the raw data for each phase, and found an effect size of ''medium'' or higher only for the difference between {\em requirements}
and {\em code}. In this case, however, 
 the overall size of the difference
{\em requirements} to {\em code} was less than 30\% (i.e. much less than we would
expect from the delayed issue effect).

@@

Nevertheless, the scale up seen in 
Figure~\ref{fig:raw} 
is much less
than that suggested by the DIE literature. Returning to
the {\em before/unit test} result: 2.56 is far smaller
than the scale ups seen in Figure~1. In fact, nowhere in 
our results do we see any exponential increase.

\newpage\subsubsection{Comment 2}
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit exponentially more
difficult issues after deployment.

That evidence is as follows.
Prior literature reports an  exponential increase in difficulty in resolving issues {\em during
development}.
For example:
\bi
\item  From requirements to acceptance testing
\fig{b81} reports a 40-fold increase;
\item
 \fig{cost-to-fix} reports a 100-fold increase
(for the larger projects).
\item
The projects surveyed in \fig{steck} report changes of a  \newline
\{37,40,40,50,51,70,75,250\}-fold increase in pre-deployment period when the software was being developed.
\ei
Any manager watching  these intra-development trends would be well-justified
to anticipate that  the difficulty in resolving issues  will get much worse. Such managers
would therefore expect an exponentially harder time resolving issues during deployment.

On the other hand, a remarkable feature of Figure~\ref{fig:raw} is the {\em absence}
of an exponential delay effect during development.
In our data, the greatest increase in difficulty in resolving requirements issues was the 1.31 scale-up seen in systems testing--
which is far smaller than the 37 to 250-fold increases reported in \fig{b81}, \fig{cost-to-fix},
and \fig{steck}.  That is, from our  data, we would not anticipate exponentially greater difficulty in resolving requirements issues after
deployment. 
 \subsubsection{Comment 3}
 The above specific statements are related just to requirements. Turning now
 to issues created in all phases in Figure~\ref{fig:raw},  we see that
 this data does not follow the exponential increase predicted
by the delayed issue effect:
\bi
\item There is much ``up and down'' in the chart, but no steady exponential growth such
as that see in, for example,  \fig{b81}.
\item The time in acceptance  testing is particularly low suggesting that few defects survived into testing
(and hence, there were few instances where code required massive rework).
\ei
%In summary:
%\bi
%\item
%The delayed issue effect has been reported in around a dozen papers, at least half of which date back to the 1970s. 
%\item
%But there is no trace of that  effect   in hundreds of software projects built  since 2006.
%\ei
%For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. Cost-to-fix for TSP is total time spent in each review, inspection, or test-related phase / \# of defects found in phase (reported in \fig{raw}). It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

@@

\begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|rl}
          \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{99th }\\
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{percentile }\\
 Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\\\hline

B4dev&Code&57&257&257&1&\rule{2mm}{2mm} \\
B4dev&CodeInspect&72&257&153&0.60&\rule{2mm}{2mm} \\
B4dev&Test&165&257&282&1.093&\rule{2mm}{2mm} \\
B4dev&IntTest&50&257&365&1.42&\rule{2mm}{2mm} \\
B4dev&SysTest&42&257&151&0.59&\rule{2mm}{2mm} \\

\\\hline

Design&Design&37&600&600&1&\rule{2mm}{2mm} \\
Design&DesignInspect&455&600&277&0.46&\rule{0mm}{2mm} \\
Design&Code&218&600&240&0.4&\rule{0mm}{2mm} \\
Design&CodeInspect&166&600&163&0.27&\rule{0mm}{2mm} \\
Design&Test&542&600&652&1.09&\rule{2mm}{2mm} \\
Design&IntTest&67&600&277&0.46&\rule{0mm}{2mm} \\
Design&SysTest&92&600&480&0.8&\rule{2mm}{2mm} \\

\\\hline

Code&Code&126&600&600&1&\rule{2mm}{2mm} \\
Code&CodeInspect&459&600&243&0.41&\rule{0mm}{2mm} \\
Code&Test&461&600&365&0.61&\rule{2mm}{2mm} \\
Code&IntTest&348&600&7200&12&\rule{24mm}{2mm} \\
Code&SysTest&230&600&290&0.48&\rule{0mm}{2mm} \\
Code&AcceptTest&71&600&600&1&\rule{2mm}{2mm} \\
\end{tabular}
\end{center}
\caption{99th percentile scale ups.}
\label{fig:scale99}
\end{figure}

@@

In this paper, we explored   the papers and data related to the 
widely-believed {\em delayed issue effect} (that delaying the resolution of issues
exponentially 
increases the difficulty of completing that  resolution).
Several prominent SE researchers state this effect is a fundamental law of software engineering~\cite{boehm01,mcconnell01,boehm01,glass02}.
Based on a  survey  of both researchers and practitioners, we  found that
a specific form  of this effect (requirements errors are hardest to fix) is  widely-believed
held in the community.

>>>>

\section{Definitions \& Hypotheses}
This paper uses the following definitions:
\bi
\item
The {\em delayed issue effect}:   it is {\em very much}  more {\em difficult} to resolve  issues in a software project, the {\em longer} they remain.
\item
 {\em Longer} time is defined as per  Boehm'81~\cite{Boehm81}; i.e. the gap between the   phases where   issues are introduced and resolved.
\item
We say that a measure collected in phase ${1,.,i,..j}$ is 
{\em very much} more when  that
   measure at phase $j$   
   is larger than the sum of that measure in 
earlier phases $1 \le i < j$. 
\item
Issues are more {\em difficult}  
when their resolution takes more time or costs more  (e.g. needs expensive
debugging tools or the skills of expensive developers).
\ei
Note that we use  the  term ``delayed issue effect'' as generalization of the
more specific rule  ``requirements errors are hardest to fix''.
This generalization is valid since the rationale for the rule about requirements
is usually done as per McConnell~\cite{mcconnell01}; i.e. small upstream mistakes very
early in the system can cascade into huge problems later in the lifecycle.
That said, we prefer our more general term ``delayed issue effect'' since not
only might requirements errors cascade, so too might analysis errors, design errors, etc.

@@

{\bf Claim3: Delayed Issues not Harder to Resolve.}
 In our sample of 
 171 software  projects developed, we will show
{\em no evidence}  that, during development, delayed issues were very much more harder to resolve the longer they were left
 in the software.

@@

This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access
the data used to make prior conclusions.
As an example of this, \fig{steck} shows one 2004 survey that reports an 
greatly increased delayed issue  effect for 
requirements issues in eight case studies. Note that we cannot verify
{\em any} of those results since the links in the references of
that  survey are all broken (``page not found'' errors).

@@

 All in all, the empirical results of this section seem insufficient to justify
 the strength of the beliefs documented in the previous section. Adding to our doubts are  several studies that report less-than large
 increase in the difficulties associated with making the changes associated with delayed
 issues. 
Clutterbuck et al.~\cite{Clutterbuck09} studied a 5-month effort by a small-to-medium enterprise team developing a 71KLOEC web interface to a database application to implement 18 change requests-- see Figure~\ref{fig:clutterbuck} (note that these were for new and changed user requirements, not defects). Clutterbuck et al. found the cost of change to be relatively flat until the later phases, with much of the effort spent in analysis of the change requests~\cite{Clutterbuck09}. Note that in this study, the effort increased by only 60\% (see the start and end of the curve in Figure~\ref{fig:clutterbuck}).

Another example of less-than large increase in the difficulty associated with delayed issues comes from  Royce~\cite{Royce98}. He
studied  a million-line, safety-critical missile defense system
(see  Figure~\ref{fig:royce}). Design changes (including architecture changes) required approximately twice the effort of implementation and test changes, and the cost-to-fix in implementation and test phases increased slowly. Boehm~\cite{Boehm10} attributes this success to the  development process, which focused on removing architecture risk early in the development lifecycle.

@@

\subseection{Early Onset of the DIE Effect}\label{sect:earlyonset}

One feature of the above results will become very important in the subsequent discussion.  
All the   literature described above that reports the onset of  DIE  {\em prior} to delivery.
\bi
\item  
\fig{b81} reports a 40-fold increase in effort requirements to acceptance testing
\item
 \fig{cost-to-fix} reports a 100-fold increase
(for the larger projects) before the code is delivered 
\item
The projects surveyed in \fig{steck} report changes of a  
\{37,40,40,50,51,70,75,250\}-fold increase in pre-deployment period when the software was being developed.
\ei
Any manager noticing this  early onset of DIE (prior to delivery, during the initial development)
would be well-justified
in believing that  the difficulty in resolving issues  will get much worse. Such managers
would therefore expect DIE to have a marked effect, post-deployment.

We make this point since,  in the new project data presented below, there is no early onset of DIE. 
That is:
\bi
\item There is no evidence of a growing problem prior to delivery that delayed issues are becoming harder to manage
\item  Hence, there there is less evidence that DIE is a trend that will significantly and negatively effect the product,
post-delivery.
\ei

@@

\begin{figure}[!t] 
% \renewcommand{\baselinestretch}{0.8}
 \scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll|rl}   
  \multicolumn{1}{c}{~} &\multicolumn{3}{c}{Percentiles}&\multicolumn{2}{c}{Phase}\\\cline{2-6}
N & 25th & 50th & 75th &   injected & removed   & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline
 48&   14&   30&   51&Before&DesignInspect &1.00 & **********  \\
 53&   13&   28&   49&Before&CodeReview &0.93 & **********  \\
 94&   22&   45&   80&Before&CodeInspect &1.50 & ***************  \\
153&   38&   77&  130&Before&UnitTest &2.56 & **************************  \\
127&   30&   65&  128&Before&IntTest &2.16 & **********************  \\
 92&   24&   56&   89&Before&SysTest &1.86 & *******************  \\\hline

 56&   14&   35&   57&Planning&ReqtsReview &1.00 & **********  \\
 46&   11&   24&   39&Planning&DesignInspect &0.68 & *******  \\
 32&   18&   33&   84&Planning&UnitTest &0.94 & **********  \\\hline

184&   45&   92&  155&Reqts&ReqtsReview &1.00 & **********  \\
189&   46&   93&  155&Reqts&ReqtsInspect &1.00 & **********  \\
 61&   14&   33&   55&Reqts&DesignReview &0.35 & ****  \\
102&   24&   49&   83&Reqts&DesignInspect &0.52 & ******  \\
 59&   18&   38&   60&Reqts&CodeInspect &0.40 & ****  \\
 91&   22&   49&  104&Reqts&UnitTest &0.52 & ******  \\
100&   33&   67&  120&Reqts&IntTest &0.72 & ********  \\
 78&   28&  122&  237&Reqts&SysTest &1.31 & **************  \\\hline 
210&   51&  103&  166&Design&DesignReview &1.01 & ***********  \\
207&   50&  101&  156&Design&DesignInspect &1.00 & **********  \\
134&   32&   66&  116&Design&CodeReview &0.65 & *******  \\
163&   39&   79&  135&Design&CodeInspect &0.78 & ********  \\
238&   58&  119&  183&Design&UnitTest &1.17 & ************  \\
172&   42&   88&  158&Design&IntTest &0.87 & *********  \\
163&   40&   80&  161&Design&SysTest &0.79 & ********  \\
 80&   21&   45&   73&Design&AcceptTest &0.44 & *****  \\\hline

235&   57&  115&  177&Code&CodeReview &1.00 &  **********  \\
228&   56&  113&  176&Code&CodeInspect &0.99 & ********** \\
243&   59&  125&  206&Code&IntTest &1.08 &     *********** \\
191&   46&   95&  173&Code&SysTest &0.82 &     ******** \\
269&   66&  133&  206&Code&UnitTest &1.15 &    ***********  \\
130&   31&   65&  122&Code&AcceptTest &0.56 &  ******  \\
 \end{tabular}
\end{center}
\caption{Distribution of fix times seen in SEI TSP data. Left hand side bars
show scale up on median values.
}
\label{fig:raw}
\end{figure}

\begin{figure}[!t] 
 %\renewcommand{\baselinestretch}{0.8}
 \scriptsize
\begin{center}
\begin{tabular}{r|r|ll|rl}   
  \multicolumn{1}{c}{~} &\multicolumn{1}{c}{Percentiles}&\multicolumn{2}{c}{Phase}\\\cline{2-4}
N  & 99th &   injected & removed   & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline

57&257&Before&Code&1.00&*********\\
72&153&Before&CodeInspect&0.60&****** \\
165&282&Before&Test&1.01&**********\\
50&365&Before&IntTest&1.42&************** \\
42&151&Before&SysTest&0.59&****** \\\hline 
37&600&Design&Design&1.00&********** \\
455&277&Design&DesignInspect&0.46&***** \\
218&240&Design&Code&0.40&**** \\
166&163&Design&CodeInspect&0.27&*** \\
542&652&Design&Test&1.09&***********\\
67&277&Design&IntTest&0.46&***** \\
92&480&Design&SysTest&0.80&******** \\\hline 
126&600&Code&Code&1.0&**********\\
459&243&Code&CodeInspect&0.4&**** \\
461&365&Code&Test&0.6&****** \\
230&290&Code&SysTest&0.5&***** \\
71&600&Code&AcceptTest&1.0&*********\\
348&7200&Code&IntTest&12.0&************************************\\
  &    &    &       &    & ************************************\\
  &    &    &       &    & ************************************\\

\end{tabular}
\end{center}
\caption{99th percentile scale ups.}
\label{fig:scale99}
\end{figure}

@@

In that figure, the data is split out according to issues that were fixed in phase $Y$ after
being introduced in phase $X$. The data is sub-divided into tables according to $X$;
i.e. according to {\em before, planning, requirements, design} or {\em  code}. 
The stars on the right-hand-side of that figure show the scale ups associated with the phase removal stages,
expressed as a ratio of the first item in that block (which, by definition, is 1.00 for
that first item).
If  this data exhibited a delayed issue effect,  we would expect those bars in each sub-division to exhibit an
very large increased.

\subsubsection{Comment 1}
Figure~\ref{fig:raw}
 data does not support that   
{\em  requirements errors are the most expensive to fix}.
Note that   median time spent fixing errors from {\em requirements} across the life cycle was 92, 93,  49, 33, 39, 49, 67, 122 minutes\footnote{If this time seems long, then recall
that it includes the time required  to (a)~collect data and realize there is an error;
(b)~prepare a fix;  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).}.
This is:
\bi
\item
Clearly longer than median times seen {\em Before } and in {\em Planning}
(30,28,45,77,65,56,24,35,33);
\item
Somewhat shorter, but still not very much larger,  than the median times see after {\em requirements} in {\em design}
(103,101, 79,66, 119,88, 80,45) and in {\em code} (65,115, 113,125, 95,133). 
\ei
For completeness, we ran an A12 test comparing the raw data for each phase, and found an effect size of ''medium'' or higher only for the difference between {\em requirements}
and {\em code}. In this case, however, 
 the overall size of the difference
{\em requirements} to {\em code} was less than 30\% (i.e. much less than we would
expect from the delayed issue effect).

@@

Nevertheless, the scale up seen in 
Figure~\ref{fig:raw} 
is much less
than that suggested by the DIE literature. Returning to
the {\em before/unit test} result: 2.56 is far smaller
than the scale ups seen in Figure~1. In fact, nowhere in 
our results do we see the kind of very large increases reported in the papers
documenting DIE.

\newpage\subsubsection{Comment 2}
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit very much more
difficulty in resolving issues after deployment.

As mentioned in \tion{earlyonset}, every other
paper reporting DIE also reported  early onset of DIE
within the current development. A remarkable feature of Figure~\ref{fig:raw} is the {\em absence}
of a very large DIE   effect during development.
In our data, the greatest increase in difficulty in resolving requirements issues was the 1.31 scale-up seen in systems testing--
which is far smaller than the 37 to 250-fold increases reported in \fig{b81}, \fig{cost-to-fix},
and \fig{steck}.

That is, from our  data, we would not anticipate very much larger  difficulty in resolving requirements issues after
deployment. 
 \subsubsection{Comment 3}
 The above specific statements are related just to requirements. Turning now
 to issues created in all phases in Figure~\ref{fig:raw},  we see that
 this data does not exhibit the very large   increases predicted
by the delayed issue effect:
\bi
\item There is much ``up and down'' in the chart, but no steady very large growth such
as that see in, for example,  \fig{b81}.
\item The time in acceptance  testing is particularly low suggesting that few defects survived into testing
(and hence, there were few instances where code required massive rework).
\ei
%In summary:
%\bi
%\item
%The delayed issue effect has been reported in around a dozen papers, at least half of which date back to the 1970s. 
%\item
%But there is no trace of that  effect   in hundreds of software projects built  since 2006.
%\ei
%For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. Cost-to-fix for TSP is total time spent in each review, inspection, or test-related phase / \# of defects found in phase (reported in \fig{raw}). It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

@@

@@

In this paper, we explored   the papers and data related to the 
widely-believed {\em delayed issue effect} (that delaying the resolution of issues
very much 
increases the difficulty of completing that  resolution).
Several prominent SE researchers state this effect is a fundamental law of software engineering~\cite{boehm01,mcconnell01,boehm01,glass02}.
Based on a  survey  of both researchers and practitioners, we  found that
a specific form  of this effect (requirements errors are hardest to fix) is  widely-believed
held in the community.
